Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-07-13 paper/main.tex
fixed most reference issues

\usepackage{xcolor}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{subfig} %ieee does not like subfigure
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{axiom}{Axiom}
\newtheorem{prop}{Proposition} %math?
\usepackage[switch]{lineno}
\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt, 
                   frame=lines,
                   autogobble,
                   framesep=3mm, 
                   breaklines=True} 
% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{tabulary}
% IEEEtran contains the IEEEeqnarray family of commands

@@

% *** PDF, URL AND HYPERLINK PACKAGES ***
\usepackage{url}

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

@@

\subsection{Data Field Structure}
\label{sec:related-work:equivariance}
Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the column types in a table, when they translate data to visual representations. The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure. Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011a}. This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}. The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}. They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e. the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner. In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes. In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}. A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

\subsection{Structure Preservation In Software}
\label{sec:related-work:software}
Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}. While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user. Libraries that provide small buildable components \cite{wongsuphasawatNavigatingWideWorld2021}, such as components for making boxes or translating data values to colors, mostly embed their connectivity assumptions in their visual algorithms and allow the user choose almost any component, regardless of structure preservation. Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the space of choices to algorithms that are appropriate for the data. 
Buildable component libraries such as matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} support a wide variety of data types, but at the cost of visual algorithms with  parameters that are tuned to the expected data in a way that leaves the over all library feeling uncohesive. VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, but does so in a non-uniform manner. On the other hand, the topology of a table as a set of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by 
A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks. Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanaties data ImagePlot\cite{studiesCulturevisImageplot2021} plugin assumed that the input is 2D continuous. Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph like structure. These domain specific libraries tend to benefit from more cohesive apis, but at the cost of a much more limited set of visualization algorithms than the building block libraries.

@@

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A map $\pi$ is a bundle projection map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial $\dtotal\restriction_{\openset} = \openset \times \dfiber$
\end{enumerate}
\end{definition}
The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fiber:fiber} and the corresponding index, discussed in \autoref{sec:atct:fiber:base}. The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fiber:sections}

\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}
In \autoref{sec:atct:fiber:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.

\subsubsection{Topological Structure: Base Space \dbase}
\autoref{sec:atct:fiber:base}
We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset. The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}. The set \dbase\ is equipped with open subsets \openset, which are a generalization of an open interval. This means opensets contain points inside a bounded region on the topological space and excludes the points on the boundary \cite{TopologyVsTopology}. An openset that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$. Opensets are key to constructing topologies, and are used in this paper to model subsets of the data contextualized within the larger continuity those subsets belong to.

@@

\subsubsection{Data Fields: Fiber Space \dfiber}
\autoref{sec:atct:fb:fiber}
As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type. To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.

@@

\subsubsection{Data: Section \dsection} 
\autoref{sec:atct:fb:sections}
We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition. We can define these section functions locally, meaning over a specific open subset \openset\ of \dbase\ 
\begin{equation}
  \label{eq:atct:fb:sections}
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\} 
\end{equation}
such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from a point $\dbasepoint \in \openset$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point. Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$. We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$, and encapsulate multiple datasets with the same fiber and base space as different sections of the same bundle.

@@

Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category. 
\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$. The functor also acts on morphisms such that for every morphism $f: C_1 \rightarrow C_2$ in category $\mathcal{C}$, there is a compatible morphism $F(f): F(C_{1}) \rightarrow F(C_2)$. 
A functor must satisfy the properties 
\begin{itemize}
  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
  \item \textit{composition}:$F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
\end{itemize}
\end{definition}
Functors preserve all the structure of the categories, namely the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}. Specifically, functors allow us to express which structure must be preserved when an object is transformed from one category to another.

@@

The presheaf preserves continuity by stating that given a morphism $\iota$ between two objects of $\mathcal{\dbase}$ 
\begin{equation*}
  \iota: \openset_1 \rightarrow \openset_2
\end{equation*}
there must be a corresponding restriction morphiism $\iota^*$ between the sets of sections 
\begin{equation*}
  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
\end{equation*}
which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$. The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects. The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container mist ensure that sections can be interpolated over any subset that the container can provide.

@@

\section{Artist: Data to Graphic}
\label{sec:artist}
In this work we propose that visualization libraries are implementing transformations from data sheaf to graphic sheaf. We call these subset of functions the artist:
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc}
\end{align}
 The artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors; therefore they are natural transformations.
 \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
\begin{itemize}
  \item \textit{data}: morphism $F(c) \xrightarrow{\alpha_{c}} G(c)$ for each object $c \in \mathcal{C}$
  \item \textit{property} when $f: c_1 \rightarrow c_2$ is a morphism in $\mathcal{C}$, the components of the natural transform commute $G(f) \circ \alpha_{c_1} = \alpha_{c_2} \circ F(f)$  
\end{itemize}
such that $\alpha = (\alpha_{c})_{c\in\mathcal{C}}$ is the set of all natural transformation components $\alpha_{c}$.\cite{bradleyWhatNaturalTransformation}
 \end{definition}
 This means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}. As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
\begin{equation}
  \label{eq:artist:sheaf:base}
    \begin{tikzcd}
      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc} 
      \end{tikzcd}
\end{equation}
take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb. As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc} 
    \arrow[dd, "\iota^*"', color=set] 
    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  & 
    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc} 
    \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] & 
    \cgamma{\dbasec_{2}}{\dtotalc} 
    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  & 
    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. Because the artist is a functor of sheaves, the artist is expected to tra nslate the data continuity to graphic continuity such that the connectivity of subsets is preserved. This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}

The output of an artist \vartist\ is a restricted subset of graphic sections
\begin{equation}
  \label{eq:artist:output}
  \imartist{\gbasec}{\gtotalc} \coloneqq\\ 
  \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\; 
  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
\end{equation} 
that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}. We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving. For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.

@@

\subsubsection{Addition: +}
\label{sec:artist:addition}
As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap. When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more records to a dataset. We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset

@@

\subsection{Implementing the Artist: $\vartistc=\vmarkc\circ\vchannelc$}

>>>>

\usepackage{xcolor}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{axiom}{Axiom}
\newtheorem{prop}{Proposition} %math?
\usepackage[switch]{lineno}
\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt, 
                   frame=lines,
                   autogobble,
                   framesep=3mm, 
                   breaklines=True} 
% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{tabulary}
% IEEEtran contains the IEEEeqnarray family of commands

@@

% *** PDF, URL AND HYPERLINK PACKAGES ***
\usepackage{url}
% \usepackage[breaklinks=true]{hyperref} breaks compilation?
% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

@@

\subsection{Data Field Structure}
\label{sec:related-work:equivariance}
Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the column types in a table, when they translate data to visual representations. The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure. Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011}. This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}. The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}. They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e. the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner. In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist} includes the correspondence between data and visual changes. In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}. A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

\subsection{Structure Preservation In Software}
\label{sec:related-work:software}
Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}. While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user. Libraries that provide small buildable components \cite{wongsuphasawatNavigatingWideWorld2021}, such as components for making boxes or translating data values to colors, mostly embed their connectivity assumptions in their visual algorithms and allow the user choose almost any component, regardless of structure preservation. Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the space of choices to algorithms that are appropriate for the data. 
Buildable component libraries such as matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} support a wide variety of data types, but at the cost of visual algorithms with  parameters that are tuned to the expected data in a way that leaves the over all library feeling uncohesive. VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, but does so in a non-uniform manner. On the other hand, the topology of a table as a set of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by 
A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks. Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanaties data ImagePlot\cite{studiesCulturevisImageplot2021} plugin assumed that the input is 2D continuous. Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph like structure. These domain specific libraries tend to benefit from more cohesive apis, but at the cost of a much more limited set of visualization algorithms than the building block libraries.

@@

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A map $\pi$ is a bundle projection map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial $\dtotal\restriction_{\openset} = \openset \times \dfiber$
\end{enumerate}
\end{definition}
The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fb:fiber} and the corresponding index, discussed in \autoref{sec:atct:fb:base}. The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fb:sections}

\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}
In \autoref{sec:atct:fb:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.

\subsubsection{Topological Structure: Base Space \dbase}
\label{sec:atct:fb:base}
We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset. The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}. The set \dbase\ is equipped with open subsets \openset, which are a generalization of an open interval. This means opensets contain points inside a bounded region on the topological space and excludes the points on the boundary \cite{TopologyVsTopology}. An openset that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$. Opensets are key to constructing topologies, and are used in this paper to model subsets of the data contextualized within the larger continuity those subsets belong to.

@@

\subsubsection{Data Fields: Fiber Space \dfiber}
\label{sec:atct:fb:fiber}
As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type. To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.

@@

\subsubsection{Data: Section} 
\label{sec:atct:fb:sections}
We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition. We can define these section functions locally, meaning over a specific open subset \openset\ of \dbase\ 
\begin{equation}
  \label{eq:atct:fb:sections}
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\} 
\end{equation}
such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from a point $\dbasepoint \in \openset$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point. Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$. We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$, and encapsulate multiple datasets with the same fiber and base space as different sections of the same bundle.

@@

Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{bradleyWhatFunctorDefinitions} that preserve the morphisms in each category. 
\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$. The functor also acts on morphisms such that for every morphism $f: C_1 \rightarrow C_2$ in category $\mathcal{C}$, there is a compatible morphism $F(f): F(C_{1}) \rightarrow F(C_2)$. 
A functor must satisfy the properties 
\begin{itemize}
  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
  \item \textit{composition}:$F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
\end{itemize}
\end{definition}
Functors preserve all the structure of the categories, namely the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}. Specifically, functors allow us to express which structure must be preserved when an object is transformed from one category to another.

@@

The presheaf preserves continuity by stating that given a morphism $\iota$ between two objects of $\mathcal{\dbase}$ 
\begin{equation*}
  \label{eq:atct:sheaf:inclusion}
  \iota: \openset_1 \rightarrow \openset_2
\end{equation*}
there must be a corresponding restriction morphiism $\iota^*$ between the sets of sections 
\begin{equation*}
  \label{eq:atct:sheaf:restriction}
  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
\end{equation*}
which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$. The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects. The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container mist ensure that sections can be interpolated over any subset that the container can provide.

@@

\section{Artist: Data to Graphic}
\label{sec:artist}
In this work we propose that visualization libraries are implementing transformations from data sheaf to graphic sheaf. We call these subset of functions the artist:
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc}
\end{align}
 The artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors; therefore they are natural transformations.
 \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
\begin{itemize}
  \item \textit{data}: morphism $F(c) \xrightarrow{\alpha_{c}} G(c)$ for each object $c \in \mathcal{C}$
  \item \textit{property} when $f: c_1 \rightarrow c_2$ is a morphism in $\mathcal{C}$, the components of the natural transform commute $G(f) \circ \alpha_{c_1} = \alpha_{c_2} \circ F(f)$  
\end{itemize}
such that $\alpha = (\alpha_{c})_{c\in\mathcal{C}}$ is the set of all natural transformation components $\alpha_{c}$.\cite{bradleyWhatNaturalTransformation}
 \end{definition}
 This means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}. As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
\begin{equation}
  \label{eq:artist:sheaf:base}
    \begin{tikzcd}
      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc} 
      \end{tikzcd}
\end{equation}
take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb. As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and \autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc} 
    \arrow[dd, "\iota^*"', color=set] 
    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  & 
    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc} 
    \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] & 
    \cgamma{\dbasec_{2}}{\dtotalc} 
    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  & 
    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform:inclusions} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. Because the artist is a functor of sheaves, the artist is expected to tra nslate the data continuity to graphic continuity such that the connectivity of subsets is preserved. This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}

The output of an artist \vartist\ is a restricted subset of graphic sections
\begin{equation}
  \label{eq:artist:output}
  \imartist{\gbasec}{\gtotalc} \coloneqq\\ 
  \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\; 
  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
\end{equation} 
that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariant:artist}. We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving. For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.

@@

\subsubsection{Addition}
\label{sec:artist:addition}
As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap. When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more records to a dataset. We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset

@@

\subsection{Implementing the Artist}
